---
title: "03-PBC_MFaivre_correction"
output: html_document
---
#http://www.uio.no/studier/emner/matnat/math/STK4080/h14/r-trial-project.txt

# Detailed commands to the trial project
# ======================================

# This only gives the R commands for the trial project. 
#It is not a draft solution of the project.
# Some comments on how to write a project report was given at the lectures.

# Read the data into R and attach the survival library

#load libraries
```{r}
library(survival)
library(survivalROC)
library(asaur)
library(ggplot2)
library(glmnet)
library(MASS)
```


#load ... dataset
```{r}
mayoBiliary <- pbc
head(mayoBiliary)
names(mayoBiliary)
summary(mayoBiliary)
print('str(mayoBiliary)')
str(mayoBiliary)

NA.hepato <- is.na(mayoBiliary$hepato)
# print(NA.hepato)
###dataHouse$bin.medv[is.na(dataHouse$bin.medv)] = nb_levels
missing.hepato = mayoBiliary$hepato[NA.hepato]
# print(mayoBiliary$status[NA.hepato]) ##values 0,1 or 2 met
```
##Data cleaning: filter out <NA> records
```{r}

```
#//////////////////// EXPLORATORY (READ DATA) ////////////////////
#=====================================================================================
#=== 1) Exploratory analysis : overall survival curve with right-censored status  ====
#=====================================================================================
#=== 1.a) summary
```{r}
summary(mayoBiliary)

colnames <-names(mayoBiliary) 
nb_obs   <-nrow(mayoBiliary)
nb_vars  <-ncol(mayoBiliary)
print(nb_vars)
print(nb_obs)
```
#=== 1.b) factor 
```{r}
mayoBiliary$sex <- factor(mayoBiliary$sex)

```


#//////////////////// MISSING VALUES AND OUTLIERS ////////////////////

#================================================================================================
#=== 2. Data cleaning : replace missing data by imputation ; detection and filter-out outliers===
#================================================================================================
'''
=== 2.a) address missing values!
cannot just filter out missing which would result in data-loss
How can we trt missing data during survival data analysis?
The cook's distance for each observation i measures the change in Y Y^ (fitted Y) for all observations with and without the presence of observation i, so we know how much the observation i impacted the fitted values.
http://publish.illinois.edu/spencer-guerrero/2014/12/11/2-dealing-with-missing-data-in-r-omit-approx-or-spline-part-1/
https://www.r-bloggers.com/imputing-missing-data-with-r-mice-package/
'''

#////////////// MORE DATA PROCESSING FOR ANALYSIS //////////////////
#== 2.b STATUS : exclude transplant and as.logical
```{r}
# STATUS : exclude transplant from Dataset
mayoBiliary <- subset(mayoBiliary, status != 1) #exlcude transplant cases
#####
mayoBiliary <- transform(mayoBiliary, status = as.logical(status)) ##DO NOT FORGET THIS OR CENSORED DATA 
#ARE FILTERED OUT !!!!
```
#=== 2.c TREATMENT
# trtMENT : NA-> 0 & filter out 'not randomized'
```{r}
mayoBiliary$trt[is.na(mayoBiliary$trt)] = 0   #'not randomized' flag
mayoBiliary <- subset(mayoBiliary, trt != 0)  #exlcude 'not randomized'
```


#=== . biligroup
```{r}
thresholds = quantile(mayoBiliary$bili)
print('thresholds')
print(thresholds)

# mayoBiliary$biligroup=cut(cirrhosis$prot,breaks=c(0,49,69,89,150), labels=1:4)
mayoBiliary$biligroup=cut(mayoBiliary$bili,breaks=thresholds, labels=1:4)

str(mayoBiliary$biligroup)
```

#=== 2.d stats on missing data
```{r}
###install.packages("VIM")
library(VIM)

# need to address NA values !
n <- names(mayoBiliary)
formula1 <- as.formula(paste("time ~", paste(n[!n %in% c("protime","id")], collapse = " + ")))
mod <- lm(formula1, data=mayoBiliary)
cooksd <- cooks.distance(mod)
# print(cooksd)

# Dataset 1 : filter all records with NA rows
PBC_noNA = na.omit(mayoBiliary)
# print(dim(PBC_noNA))
  
# Dataset 2 : Proceed Imputation with mean value (as a first approx.) per each row
#Imputing the missing data
#The mice() function takes care of the imputing process
###install.packages("mice")
library(mice)
temp_NAimput <- mice(mayoBiliary, m=5, maxit=50, meth='pmm', seed=500)
PBC_NAimput  <- complete(temp_NAimput,1)
# summary(PBC_NAimput)
####densityPlot(temp_NAimput)
  
# Dataset 3 : apply a row-based spline to replace NA missing values 
#PBC_NAspline = 

#mayoBiliary = PBC_noNA
mayoBiliary <- PBC_NAimput
mean.survtime <- mean(mayoBiliary$time)
# print(mean.survtime/365.24)

# TIME IN YEARS
mayoBiliary$time <- mayoBiliary$time/365.24

# summary(mayoBiliary)
```
#//////////////////// MORE EXPLORATORY PLOTS ////////////////////

#==================================================
#=== 2.e scatter plots wr survival time        ===
#==================================================
'''
```{r}
# 1. STime vs bilirubin
scatter.smooth(mayoBiliary$bili, mayoBiliary$time)

# 2. STime vs choloresterol
scatter.smooth(mayoBiliary$chol, mayoBiliary$time)

# 3. STime vs albumin
scatter.smooth(mayoBiliary$albumin, mayoBiliary$time)

# 4. STime vs copper
scatter.smooth(mayoBiliary$copper, mayoBiliary$time)

# 5. STime vs platelet
scatter.smooth(mayoBiliary$platelet, mayoBiliary$time)

# 6. STime vs alk.phos
scatter.smooth(mayoBiliary$alk.phos, mayoBiliary$time)
```

#=== 2.f Log-linearity of the numeric covariates Beta(t) per covariate in he model Cox has been checked along the way using splines 
# [both for the unvariate and mulivariate Cox models (the latter commands not given here)]

# We also need to check for proportional hazards:
# We find an interaction between trtment and ascites, and between sex and age.
# To ease the interpretation of the interaction between sex and the numeric covariate age, it is useful
# to center age by subtracting 60 years (which is close to the mean age) 
```{r}
mayoBiliary$cage10=(mayoBiliary$age-60)/10
cox.final=coxph(Surv(time,status)~trt+sex+ascites+cage10+bili+trt:ascites+sex:cage10, data=mayoBiliary)
print('summary(cox.final)')
summary(cox.final)

cox.test=cox.zph(cox.final,transform='log')
print(cox.test)
par(mfrow=c(2,2))
plot(cox.test)
```


#3. plot overall KME survival curve with censored data indicated
#===============================================================
```{r}
plot(survfit(Surv(time, status) ~ 1, data=mayoBiliary), main='Overall KME survival time')

```

#/////////////////// UNVARIATE KAPLAN-MEIER & NELSON-AALEN ////////////////

#4. NON-PARAMETRIC METHOD UNIVARIATE
#==========================================================

# 4.a Simple univariate analyses for one covariate at a time
#==========================================================
 
# Treatment
# ---------
# Nelson-Aalen and Kaplan-Meier plots stratified:
```{r}
par(mfrow=c(1,2))
fit.trt.naa=coxph(Surv(time,status)~strata(trt),data=mayoBiliary)
surv.trt.naa=survfit(fit.trt.naa)

plot(surv.trt.naa,fun="cumhaz", mark.time=T , col=1:3,lty=1:3, 
     xlab="Years since randomization",ylab="Cumulative hazard", main="Treatment")
legend("topleft",legend=c("D-penicill","Placebo"),col=1:3,lty=1:3)

fit.trt.km=survfit(Surv(time,status)~trt,data=mayoBiliary, conf.type="plain")

plot(fit.trt.km, mark.time=T, col=1:3, lty=1:3, 
     xlab="Years since randomization", ylab="Survival",main="Treatment")
legend("bottomleft",legend=c("D-penicill","Placebo"),col=1:3,lty=1:3)

# Estimates of five years survival probabilities with "plain" confidence intervals 
# (alternatively we could have used the option "log-log")
print('summary(fit.trt.km,time=5)')
summary(fit.trt.km,time=5) 

# Estimates of median survival time 
print('fit.trt.km')
print(fit.trt.km)
#
# Alternatively, we may obtain the quartiles by the command: 
print('quantile(fit.trt.km)')
quantile(fit.trt.km)

# Log-rank test:
print('survdiff(Surv(time,status)~trt,data=mayoBiliary)')
survdiff(Surv(time,status)~trt,data=mayoBiliary)
```

# Sex
# ---

# Nelson-Aalen and Kaplan-Meier plots stratified:
```{r}
par(mfrow=c(1,2))
fit.sex.naa=coxph(Surv(time,status)~strata(sex),data=mayoBiliary)
surv.sex.naa=survfit(fit.sex.naa)
plot(surv.sex.naa,fun="cumhaz", mark.time=T ,col=1:2, lty=1:2, 
     xlab="Years since randomization",ylab="Cumulative hazard", main="Sex")
legend("topleft",legend=c("Male","Female"), col=1:2, lty=1:2)
fit.sex.km=survfit(Surv(time,status)~sex,data=mayoBiliary, conf.type="plain")
plot(fit.sex.km, mark.time=T, col=1:2, lty=1:2, 
    xlab="Years since randomization", ylab="Survival",main="Sex")
legend("topright",legend=c("Male","Female"), col=1:2,lty=1:2)

# plot with error bars


# Estimates of five years survival probabilities with "plain" confidence intervals 
print('summary(fit.sex.km,time=5)')
summary(fit.sex.km,time=5) 

# Estimates of median survival time 
print('fit.sex.km')
print(fit.sex.km)

# Log-rank test:
print('survdiff(Surv(time,status)~sex,data=mayoBiliary)')
survdiff(Surv(time,status)~sex,data=mayoBiliary)
```

# Ascites
# -------
```{r}
# Nelson-Aalen and Kaplan-Meier plots:
par(mfrow=c(1,2))
fit.asc.naa=coxph(Surv(time,status)~strata(ascites),data=mayoBiliary)
surv.asc.naa=survfit(fit.asc.naa)
plot(surv.asc.naa,fun="cumhaz", mark.time=T ,lty=1:3, 
     xlab="Years since randomization",ylab="Cumulative hazard", main="Ascites")
legend("topleft",legend=c("None","Slight","Marked"),lty=1:3)
fit.asc.km=survfit(Surv(time,status)~ascites,data=mayoBiliary, conf.type="plain")
plot(fit.asc.km, mark.time=T, lty=1:3, 
     xlab="Years since randomization", ylab="Survival",main="Ascites")
legend("topright",legend=c("None","Slight","Marked"),lty=1:3)

# Estimates of five years survival probabilities with "plain" confidence intervals 
print('summary(fit.asc.km,time=5)')
summary(fit.asc.km,time=5) 

# Estimates of median survival time 
print('fit.asc.km')
print(fit.asc.km)

# Log-rank test:
print('survdiff(Surv(time,status)~ascites,data=mayoBiliary)')
survdiff(Surv(time,status)~ascites,data=mayoBiliary)
```


# Age
# ---
```{r}
# First we create a categorical variable for age group:
mayoBiliary$agegroup=cut(mayoBiliary$age,breaks=c(0,49,59,69,100), labels=1:4)

# Nelson-Aalen and Kaplan-Meier plots:
par(mfrow=c(1,2))
fit.age.naa=coxph(Surv(time,status)~strata(agegroup),data=mayoBiliary)
surv.age.naa=survfit(fit.age.naa)
plot(surv.age.naa,fun="cumhaz", mark.time=F ,lty=1:4, 
     xlab="Years since randomization",ylab="Cumulative hazard", main="Age")
legend("topleft",legend=c("Below 50","50-59","60-69","70 and above"),lty=1:4)
#
fit.age.km=survfit(Surv(time,status)~agegroup,data=mayoBiliary, conf.type="plain")
plot(fit.age.km, mark.time=F, lty=1:4, 
     xlab="Years since randomization", ylab="Survival",main="Age")
legend("bottomleft",legend=c("Below 50","50-59","60-69","70 and above"),lty=1:4)

# Estimates of five years survival probabilities with "plain" confidence intervals 
summary(fit.age.km,time=5) 

# Estimates of median survival time 
print(fit.age.km)

# Log-rank test:
survdiff(Surv(time,status)~agegroup,data=mayoBiliary)
```
# Edema categorical univariate
# -------
```{r}
# Nelson-Aalen and Kaplan-Meier plots:
par(mfrow=c(1,2))
fit.edema.naa=coxph(Surv(time,status)~strata(edema),data=mayoBiliary)
surv.edema.naa=survfit(fit.asc.naa)

plot(surv.edema.naa,fun="cumhaz", mark.time=T ,lty=1:3, 
     xlab="Years since randomization",ylab="Cumulative hazard", main="Edema")
legend("topleft",legend=c("No edema","untreated or success","edema"),lty=1:3)
fit.edema.km=survfit(Surv(time,status) ~ edema,data=mayoBiliary, conf.type="plain")

plot(fit.edema.km, mark.time=T, lty=1:3, 
     xlab="Years since randomization", ylab="Survival",main="Edema")
legend("topright",legend=c("No edema","untreated or success","edema"),lty=1:3)

# Estimates of five years survival probabilities with "plain" confidence intervals 
summary(fit.edema.km,time=5) 

# Estimates of median survival time 
print(fit.edema.km)

# Log-rank test:
survdiff(Surv(time,status)~edema,data=mayoBiliary)
```
#hepato
# -------
```{r}
# Nelson-Aalen and Kaplan-Meier plots:
par(mfrow=c(1,2))
fit.hepato.naa=coxph(Surv(time,status)~strata(hepato),data=mayoBiliary)
surv.hepato.naa=survfit(fit.asc.naa)
plot(surv.hepato.naa,fun="cumhaz", mark.time=T ,lty=1:2, 
     xlab="Years since randomization",ylab="Cumulative hazard", main="hepato")
legend("topleft",legend=c("No hepato","hepato"),lty=1:2)
fit.hepato.km=survfit(Surv(time,status) ~ hepato,data=mayoBiliary, conf.type="plain")
plot(fit.hepato.km, mark.time=T, lty=1:2, 
     xlab="Years since randomization", ylab="Survival",main="hepato")
legend("topright",legend=c("No hepato","hepato"),lty=1:2)

# Estimates of five years survival probabilities with "plain" confidence intervals 
summary(fit.hepato.km,time=5) 

# Estimates of median survival time 
print(fit.hepato.km)

# Log-rank test:
survdiff(Surv(time,status)~hepato,data=mayoBiliary)
```



#////////////////////// UNIVARIATE COX-REGRESSION ///////////////

# 5. Univariate Cox regressions
# =============================

# treatment 
# ---------
```{r}
cox.trt=coxph(Surv(time,status)~trt,data=mayoBiliary)
print("summary(cox.trt)")
summary(cox.trt)
pred.trt = survfit(cox.trt, data=mayoBiliary, type="aalen")
```


# Sex 
# ---
```{r}
cox.sex=coxph(Surv(time,status)~sex,data=mayoBiliary)
print('summary(cox.sex)')
summary(cox.sex)

pred.sex = survfit(cox.sex, data=mayoBiliary, type="aalen")

par(mfrow=c(1,2))
plot(pred.trt, col=1:2, fun = "cloglog")
title(main="PH Testing for covariates treatment",cex.main=0.8,xlab="Time(year)",ylab="S(t)")

plot(pred.sex, col=1:2, fun = "cloglog")
title(main="PH Testing for covariates sex",cex.main=0.8,xlab="Time(year)",ylab="S(t)")
```

#Ascites:
# -------
```{r}
cox.asc=coxph(Surv(time,status)~ascites,data=mayoBiliary)
print('summary(cox.asc)')
summary(cox.asc)
```

# Age
# ---
```{r}
# For the numeric covariates age and prothrombin index, we need to decide how thay should be coded  
# (as given on the data file, or suitably transformed, or grouped).

# To see how age should be coded, we fit a model using a spline for age:
cox.psage=coxph(Surv(time,status)~pspline(age),data=mayoBiliary)
print('cox.psage')
print(cox.psage)
par(mfrow=c(1,1))
termplot(cox.psage,se=T)

# Both the plot (from the templot-command) and the test (from the print-command) show that 
# it is reasonable to assume a log-linear effect of age

# We will therefore fit a Cox model using age as a numeric covariate.
# It may be sensible to report the effect of age per 10 years, so we define a new covariate 
# where age is given per 10 years and fit a Cox model with this covariate
mayoBiliary$age10= mayoBiliary$age/10
cox.age10=coxph(Surv(time,status)~age10,data=mayoBiliary)
print('cox.age10')
summary(cox.age10)
```

#6. KME PH univariate
#====================

#=== 6.a Plotting the CUMULATIVE HAZARD figures for Gender
```{r}
fit.KM.sex <- survfit(Surv(time, status) ~ factor(sex), data=mayoBiliary)
plot(fit.KM.sex$time, log(-log(fit.KM.sex$surv)), col=1:2, type="s",xlab ="Time(year)", ylab = "log-log S(t)", main = "Proportional hazard testing for Gender", lwd=1.4)
legend("bottomright", col = 1:2, lty = 1:2, legend = c("m","f"),bty="n")
```

#7. ROC "low-risk, high-risk" for 312 formal study participants from the matching 'mayo' dataset
#===============================================================================================
312 formal study participants in common with mayoBiliary dataset (+106 eligible nonenrolled subjects)
```{r}
library(survival)
library(survivalROC)

data(mayo)
##plot(survfit(Surv(time/365.25, censor) ~ 1, data=mayo))

print(str(mayo))
str(mayo)

ROC.4 = survivalROC(Stime = mayo$time,
                status= mayo$censor,
                marker= mayo$mayoscore4,
                predict.time = 365.25*5,
                method="KM")

ROC.5 = survivalROC(Stime = mayo$time,
                status= mayo$censor,
                marker= mayo$mayoscore5,
                predict.time = 365.25*5,
                method="KM")

'list from 2 models'
ROC = list(mayo4 = ROC.4, mayo5 = ROC.5)

cutoff = with(ROC$mayo5, min(cut.values[FP<=0.1]))

mayo$prediction = ifelse(mayo$mayoscore5 <= cutoff, "low_risk","high_risk")

'compare 2 groups : low_risk vs high_risk'
'Predict Survival Time taking into account censoring wr covariate=binary prediction{"high_risk","low_risk"}'
fit.KM = survfit(Surv(time/365.24,censor)~prediction, data=mayo)

plot(fit.KM,col=1:2,mark.time=T,lty=1:2, 
     xlab="Years since randomization", ylab="Survival",main="Risk level")
legend("bottomleft",legend=c("High-risk","Low-risk"),col=1:2,lty=1:2)

'+ show censored data!'
'mayo$prediction = sapply(ROC, AUC)'
```


#////////////// MUTLI-VARIATE COX REGRESSION /////////////

#6. Multivariate Cox-PH
#======================
# c. Multivariate Cox regression
# ==============================
```{r}
# We then fit a Cox model with all the covariates 
# cox.all=coxph(Surv(time,status)~trt+sex+ascites+age10+bili, data=mayoBiliary)
cox.all=coxph(Surv(time,status)~trt+sex+ascites+age10+biligroup, data=mayoBiliary)
summary(cox.all)
```
# In principle it may be the case that the coding of a numeric covariate that is appropriate 
# for a univariate analysis, is not appropriate for a multivariate analysis and vice versa. 
# But this does not seem to be the case here (commands not shown)

# We then check for first order interactions between any pair of two covariates
#         ---------------------------------------------------------------------       

# treatment and sex:
```{r}
cox.trt.sex=coxph(Surv(time,status)~trt+sex+ascites+age10+ biligroup+trt:sex, data=mayoBiliary)
anova(cox.all,cox.trt.sex)
summary(cox.trt.sex)

pred.trt.sex = survfit(cox.trt.sex, data=mayoBiliary, type="aalen")

plot(pred.trt.sex, col=1:2)
title(main="PH Testing for covariates treat&sex",cex.main=0.8,xlab="Time(year)",ylab="S(t)")

plot(pred.trt.sex, col=1:2, fun = "cloglog")
title(main="PH Testing for covariates treat&sex cloglog",cex.main=0.8,xlab="Time(year)",ylab="S(t)")
```

# treatment and ascitesites:
```{r}
cox.trt.ascites=coxph(Surv(time,status)~trt+sex+ascites+age10+ biligroup+trt:ascites, data=mayoBiliary)
anova(cox.all,cox.trt.ascites)
summary(cox.trt.ascites)

pred.trt.asc = survfit(cox.trt.ascites, data=mayoBiliary, type="aalen")
```
# treatment and age:
```{r}
cox.trt.age=coxph(Surv(time,status)~trt+sex+ascites+age10+ biligroup+trt:age10, data=mayoBiliary)
anova(cox.all,cox.trt.age)
pred.trt.age = survfit(cox.trt.age, data=mayoBiliary, type="aalen")
summary(cox.trt.age)

par(mfrow=c(1,2))
plot(pred.trt.age, col=1:2)
title(main="PH Testing for covariates treat&age",cex.main=0.8,xlab="Time(year)",ylab="S(t)")

plot(pred.trt.asc, col=1:2)
title(main="PH Testing for covariates treat&ascites",cex.main=0.8,xlab="Time(year)",ylab="S(t)")

par(mfrow=c(1,2))
plot(pred.trt.age, col=1:2, fun = "cloglog")
title(main="PH Testing for covariates treat&age cloglog",cex.main=0.8,xlab="Time(year)",ylab="S(t)")

plot(pred.trt.asc, col=1:2, fun = "cloglog")
title(main="PH Testing for covariates treat&ascites cloglog",cex.main=0.8,xlab="Time(year)",ylab="S(t)")
```

# Sex and ascites:
```{r}
cox.sex.ascites=coxph(Surv(time,status)~trt+sex+ascites+age10+ biligroup+sex:ascites, data=mayoBiliary)
anova(cox.all,cox.sex.ascites)
pred.sex.asc = survfit(cox.sex.ascites, data=mayoBiliary, type="aalen")
summary(cox.sex.ascites)
```

# Sex and age:
```{r}
cox.sex.age=coxph(Surv(time,status)~trt+sex+ascites+age10+biligroup+sex:age10, data=mayoBiliary)
anova(cox.all,cox.sex.age)
pred.sex.age = survfit(cox.sex.age, data=mayoBiliary, type="aalen")
summary(cox.sex.age)

par(mfrow=c(1,2))
plot(pred.sex.age, col=1:2, fun = "cloglog")
title(main="PH Testing for covariates gender&age",cex.main=0.8,xlab="Time(year)",ylab="S(t)")

plot(pred.sex.asc, col=1:2, fun = "cloglog")
title(main="PH Testing for covariates gender&ascites",cex.main=0.8,xlab="Time(year)",ylab="S(t)")

# Plot the baseline survival function of this Cox PH model
install.packages("survminer")
library("survminer")
ggsurvplot(survfit(cox.sex.age), color = "#2E9FDF",
 ggtheme = theme_minimal(), title="Baseline survival function of Cox PH model:Surv(time,status)~trt+sex+ascites+age10+bili+sex:age10")
```

# Sex and bilirubin:
```{r}
cox.sex.bili=coxph(Surv(time,status)~trt+sex+ascites+age10+biligroup+sex:biligroup, data=mayoBiliary)
anova(cox.all,cox.sex.bili)
summary(cox.sex.bili)
```
# ascitesites and age:
```{r}
cox.ascites.age=coxph(Surv(time,status)~trt+sex+ascites+age10+bili+ascites:age10, data=mayoBiliary)
anova(cox.all,cox.ascites.age)
summary(cox.ascites.age)
```
# Ascites and prothrombin:
# For checking interaction between ascites and prothrombin group, we need to merge the two highest 
# prothrombin groups (since there is only one person with severe ascites in the highest prothrombin group)
# We find an interaction between treatment and ascites, and between sex and age.
# To ease the interpretation of the interaction between sex and the numeric covariate age, it is useful  
# center age by subtracting 60 years (which is close to the mean age) 
```{r}
mayoBiliary$cage10=(mayoBiliary$age-60)/10
cox.final=coxph(Surv(time,status)~trt+sex+ascites+cage10+biligroup+trt:ascites+sex:cage10, data=mayoBiliary)
summary(cox.final)
```

# Log-linearity of the numeric covariates has been checked along the way using splines 
# [both for the unvariate and mulivariate Cox models (the latter commands not given here)]

# We also need to check for proportional hazards:
```{r}
cox.test=cox.zph(cox.final,transform='log')
print(cox.test)
par(mfrow=c(2,2))
plot(cox.test)
```



#/////////// MACHINE LEARNING : RANDOM FOREST ///////////

#7. Random Forest with radomforestSRC package
#============================================

#for this part, we add the rm function to clean our memory because we don't 
#need the latest objects to continue our study
```{r}
rm(list = ls())
##install.packages("ggRandomForests")
library("ggRandomForests")
#library("ggplot2")
library("dplyr")
data(pbc, package="randomForestSRC")
head(pbc[is.na(pbc$treatment),], n = 50)

#Please consider a more traditional train/test split, only with the 312 complete data:
pbc2 <- pbc[!is.na(pbc$treatment), ]

smp_size <- floor(0.70 * nrow(pbc2))

## set the seed to make your partition reproductible
set.seed(123)
train_ind <- sample(seq_len(nrow(pbc2)), size = smp_size)

pbc.train <- pbc2[train_ind, ]
pbc.test  <- pbc2[-train_ind, ]
nrow(pbc.train)
## [1] 218
nrow(pbc.test)
## [1] 94
##build model
rfsrc_pbc <- rfsrc(Surv(days, status) ~ .,
                   data = pbc.train)
##plot the random survival forest
ggRFsrc <- plot(gg_rfsrc(rfsrc_pbc), alpha = 0.2) +
#scale_color_manual(values = strCol) +
theme(legend.position = "none") +
labs(y = "Survival Probability", x = "Time (Months)") +
coord_cartesian(ylim = c(-0.01, 1.01))
ggRFsrc
```

##test model - test data contains un-censored data
```{r}
test.pred.rfsrc <- predict(rfsrc_pbc, pbc.test,
na.action="na.impute") #added this so I get results for all test rows
#summary of our test model
summary(test.pred.rfsrc)
```

